Eco-evolutionary dynamics of finite populations from first principles

### 2023-02-27 ind_files/05_examples.tex
Make fourier stuff into section instead of subsection

\section{Example in one dimension: The stochastic logistic equation}
Consider the functional forms of example \ref{ex_1D_stoch_logistic}, given by
\begin{equation}
\label{ex_1D_stoch_logistic_BD_eqns}
\begin{aligned}
    b(n) &= \lambda n\\
    d(n) &= \left(\mu + (\lambda-\mu)\frac{n}{K}\right)n
\end{aligned}
\end{equation}
Here, $K$ is the system-size parameter. Introducing the new variable $x=n/K$, we obtain
\begin{align*}
    b_K(x) &= \frac{1}{K}b(n) = \frac{1}{K}\lambda Kx\\
    d_K(x) &= \frac{1}{K}d(n) = \frac{1}{K}\left(\mu + (\lambda-\mu)\frac{Kx}{K}\right)Kx
\end{align*}
Thus, we have
\begin{equation*}
    A^{\pm}(x) = b_K(x)\pm d_K(x) = x\left(\lambda \pm \left(\left(\mu + (\lambda-\mu)x\right)\right) \right)
\end{equation*}
Defining $r=\lambda-\mu$ and $v=\lambda+\mu$ and using equation \eqref{1D_SDE}, we 
see that the `mesoscopic view' of the system is given by the solution of the SDE
\begin{equation}\label{ex_1D_stoch_logistic_full_SDE}
dX_t =  rX_t(1-X_t)dt + \sqrt{\frac{X_t(v+rX_t)}{K}}dB_t
\end{equation}
From equation \eqref{1D_det_limit}, we see that the deterministic dynamics are
\begin{equation}\label{ex_1D_stoch_logistic_det_limit}
\frac{dx}{dt} = A^-(x) = rx(1-x)
\end{equation}
showing that in the infinite population limit, we
obtain the logistic equation. Letting $\alpha(t)$ be the solution of the logistic equation \eqref{ex_1D_stoch_logistic_det_limit}, We can Taylor expand $A^{\pm}(x)$ for the weak noise approximation, and we find:
\begin{align*}
A^-_1(x) &= \frac{d}{dx}(rx(1-x))\biggl{|}_{x=\alpha} = r(1 - 2\alpha(t))\\
A^+_0(x) &= \alpha(t)(v+r\alpha(t))
\end{align*}
Thus, the weak noise approximation of \ref{ex_1D_stoch_logistic_BD_eqns} is given by
\begin{equation}
    X_t = \alpha(t) + \frac{1}{\sqrt{K}}Y_t
\end{equation}
where the stochastic process $Y_t$ is an Ornstein-Uhlenbeck process given by the solution to the linear SDE
\begin{align}
    dY_t &= A^-_1(t)Y_tdt + \sqrt{A^+_0(t)}dB_t\nonumber\\
    \Rightarrow dY_t &= r(1 - 2\alpha(t))Y_tdt + \sqrt{\alpha(t)(v+r\alpha(t))}dB_t\label{ex_1D_stoch_logistic_WNA}
\end{align}

@@

\section{Example for discrete traits: Lotka-Volterra and matrix games in finite populations}

@@

\hl{can compare time series or something here if we think it is helpful. Visualization may be little hard though, because we always have $\geq 3$ dimensions ($m$, + 1 additional for time)}

\section{Example for quantitative traits}

\subsection{Interlude: Detecting clusters through Fourier analysis}

@@

\subsection{An example: The quantitative logistic equation}

>>>>

\section{An example in one dimension: The stochastic logistic equation}
Consider the functional forms of example \ref{ex_1D_stoch_logistic}, given by
\begin{equation}
\label{ex_1D_stoch_logistic_BD_eqns}
\begin{aligned}
    b(n) &= \lambda n\\
    d(n) &= \left(\mu + (\lambda-\mu)\frac{n}{K}\right)n
\end{aligned}
\end{equation}
Here, $K$ is the system-size parameter. Introducing the new variable $x=n/K$, we obtain
\begin{align*}
    b_K(x) &= \frac{1}{K}b(n) = \frac{1}{K}\lambda Kx\\
    d_K(x) &= \frac{1}{K}d(n) = \frac{1}{K}\left(\mu + (\lambda-\mu)\frac{Kx}{K}\right)Kx
\end{align*}
Thus, we have
\begin{equation*}
    A^{\pm}(x) = b_K(x)\pm d_K(x) = x\left(\lambda \pm \left(\left(\mu + (\lambda-\mu)x\right)\right) \right)
\end{equation*}
Defining $r=\lambda-\mu$ and $v=\lambda+\mu$ and using equation \eqref{1D_SDE}, we 
see that the `mesoscopic view' of the system is given by the solution of the SDE
\begin{equation}\label{ex_1D_stoch_logistic_full_SDE}
dX_t =  rX_t(1-X_t)dt + \sqrt{\frac{X_t(v+rX_t)}{K}}dB_t
\end{equation}
From equation \eqref{1D_det_limit}, we see that the deterministic dynamics are
\begin{equation}\label{ex_1D_stoch_logistic_det_limit}
\frac{dx}{dt} = A^-(x) = rx(1-x)
\end{equation}
showing that in the infinite population limit, we
obtain the logistic equation. Letting $\alpha(t)$ be the solution of the logistic equation \eqref{ex_1D_stoch_logistic_det_limit}, We can Taylor expand $A^{\pm}(x)$ for the weak noise approximation, and we find:
\begin{align*}
A^-_1(x) &= \frac{d}{dx}(rx(1-x))\biggl{|}_{x=\alpha} = r(1 - 2\alpha(t))\\
A^+_0(x) &= \alpha(t)(v+r\alpha(t))
\end{align*}
Thus, the weak noise approximation of \ref{ex_1D_stoch_logistic_BD_eqns} is given by
\begin{equation}
    X_t = \alpha(t) + \frac{1}{\sqrt{K}}Y_t
\end{equation}
where the stochastic process $Y_t$ is an Ornstein-Uhlenbeck process given by the solution to the linear SDE
\begin{align}
    dY_t &= A^-_1(t)Y_tdt + \sqrt{A^+_0(t)}dB_t\nonumber\\
    \Rightarrow dY_t &= r(1 - 2\alpha(t))Y_tdt + \sqrt{\alpha(t)(v+r\alpha(t))}dB_t\label{ex_1D_stoch_logistic_WNA}
\end{align}

@@

\section{An example for discrete traits: Lotka-Volterra and matrix games in finite populations}

@@

\section{Interlude: Detecting modes in quantitative trait distribution through Fourier analysis}

@@

\section{An example for quantitative traits: The quantitative logistic equation}
